Exploring Prosper Loan Data by Pyari Singh K
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(lubridate)
library(dplyr)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
prosperLoanData <- read.csv("prosperLoanData.csv")


```



> **Tip**: For the explore and summarise data project, the propserLoanData 
data set is what I chose. The data set has 81 columns & 113937 observations.  


```{r}
#converting the ListingCategory to factor variable & associating them 
#for labeling purposes
prosperLoanData$ListingCategory..numeric. <- 
  factor(prosperLoanData$ListingCategory..numeric.)

cats <- c('0 - Not Available', '1 - Debt Consolidation', 
          '2 - Home Improvement', '3 - Business', '4 - Personal Loan', 
          '5 - Student Use', '6 - Auto', '7- Other', '8 - Baby&Adoption', 
          '9 - Boat', '10 - Cosmetic Procedure', '11 - Engagement Ring', 
          '12 - Green Loans', '13 - Household Expenses', 
          '14 - Large Purchases', '15 - Medical/Dental', '16 - Motorcycle', 
          '17 - RV', '18 - Taxes', '19 - Vacation', '20 - Wedding Loans')
#converting Term into a factor variable
prosperLoanData$Term <- factor(prosperLoanData$Term)

```

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, Univariate_Plots_15, message = FALSE, warning= FALSE}
ggplot(aes(x=LoanOriginalAmount),data=prosperLoanData)+geom_histogram()+ggtitle("Histogram for Loan Original Amount")
meanLoanOriginalAmount <- mean(prosperLoanData$LoanOriginalAmount)
medianLoanOriginalAmount <- median(prosperLoanData$LoanOriginalAmount)
```


Loan Original Amount has a positive skew. The mean value for the 
LoanOriginalAmount is 8337 which is greater than the median. The median for the
LoanOriginalAmount is 6500. 

```{r echo=FALSE, Univariate_Plots_16, message = FALSE, warning= FALSE}
ggplot(aes(x=MonthlyLoanPayment),data=prosperLoanData)+geom_histogram(binwidth = 50)+
  ggtitle("Histogram for Monthly Loan Payment")+coord_cartesian(xlim=c(0:1500))
monthlyLoanPaymentMean <- mean(prosperLoanData$MonthlyLoanPayment)
monthlyLoanPaymentMedian <- median(prosperLoanData$MonthlyLoanPayment) 
```

Monthly Loan Payment has a positive skew. The median is less than the mean for 
this data. The median for the monthly Loan Payment is 217 and the mean is 272. 


```{r echo=FALSE, Univariate_Plots_17, message = FALSE, warning= FALSE}
ggplot(aes(x=CreditScoreRangeLower),data=prosperLoanData)+geom_histogram(binwidth = 5)+
  ggtitle("Histogram for Credit Score Range Lower") + xlim(400, 900)


creditScoreRangeLower_na_removed<- na.omit(prosperLoanData$CreditScoreRangeLower)
creditScoreRangeLowerMean <- mean(creditScoreRangeLower_na_removed)
creditScoreRangeLowerMedian <- median(creditScoreRangeLower_na_removed) 
```


credit score range lower has a normal distribution. The mean/median in this case 
is 685.56/680


```{r echo=FALSE, Univariate_Plots_18, message = FALSE, warning= FALSE}
ggplot(aes(x=BorrowerRate),data=prosperLoanData)+geom_histogram(binwidth = 0.01)+
  ggtitle("Histogram for Borrower Rate")
BorrowerRateMean <- mean(prosperLoanData$BorrowerRate)
BorrowerRateMedian <- median(prosperLoanData$BorrowerRate) 
```


BorrowerRate also has a normal distribution and the mean/med ratio is 0.19/0.18
However, there are various spikes especially between 0.24 and 0.31.


```{r echo=FALSE, Univariate_Plots_19, message=FALSE, warning=FALSE}
ggplot(aes(x=DebtToIncomeRatio),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Debt to Income Ratio")+xlim(0,1.75)
DIRatio_na_removed<- na.omit(prosperLoanData$DebtToIncomeRatio)
DIRatioMean <- mean(DIRatio_na_removed)
DIRatioMedian <- median(DIRatio_na_removed) 
```


Debt To Income Ratio exhibits a slight positive skew with mean > median. 
Mean is 0.275 and median is 0.22. There is a strong observation here. DI ratio 
is less than 0.5 for a good percentage of pepole. 



```{r echo=FALSE, Univariate_Plots_20, warning=FALSE, message=FALSE}
ggplot(aes(x=StatedMonthlyIncome),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Stated Monthly Income")+xlim(0,20000)
#DIRatio_na_removed<- na.omit(prosperLoanData$StatedMonthlyIncome)
StatedMonthlyIncomeMean <- mean(prosperLoanData$StatedMonthlyIncome)
StatedMonthlyIncomeMedian <- median(prosperLoanData$StatedMonthlyIncome) 
```


Stated Monthly Income has a positive skew with mean > median. Mean for the 
Stated Monthly Income is 5608 and the median is 4666. 


```{r echo=FALSE, Univariate_Plots_21, warning=FALSE, message=FALSE}

prosperLoanData$MonthlyDebtTotal <- prosperLoanData$DebtToIncomeRatio * 
  prosperLoanData$StatedMonthlyIncome
ggplot(aes(x=MonthlyDebtTotal),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Total Monthly Debt")+xlim(0,7500)
TotalMonthlyDebt_na_removed<- na.omit(prosperLoanData$MonthlyDebtTotal)
TotalMonthlyDebtMean <- mean(TotalMonthlyDebt_na_removed)
TotalMonthlyDebtMedian <- median(TotalMonthlyDebt_na_removed) 
summary(prosperLoanData$MonthlyDebtTotal)
```


MonthlyDebtTotal has been added as a new column. This was derived by multiplying
the debt to Income Ratio to the stated Monthly Income. 

Total Monthly debt has a positive skew on its histogram and the mean is greater 
than the median. The mean is 1224. The median is 1061. The first quartile for 
this data is 637 and the third quartile is 1622. 

```{r echo=FALSE, Univariate_Plots_22, warning=FALSE, message=FALSE}


ggplot(aes(x=EmploymentStatusDuration),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Employment Status Duration")+xlim(0,700)
EmploymentStatusDuration_na_removed<- na.omit(prosperLoanData$EmploymentStatusDuration)
EmploymentStatusDurationMean <- mean(EmploymentStatusDuration_na_removed)
EmploymentStatusDurationMedian <- median(EmploymentStatusDuration_na_removed) 

```


The Employment Status Duration exhibits a positive skew. The mean of the data is
higher than the median. The mean is 96 and the median is 67. It is observed that
more number of employees worked for shorter durations. 



```{r echo=FALSE, Univariate_Plots_9, warning=FALSE, message=FALSE}


ggplot(aes(x=CurrentCreditLines),data=prosperLoanData)+geom_histogram(binwidth = 1)+
  ggtitle("Histogram for Current Credit Lines")+xlim(0,70)
CurrentCreditLines_na_removed<- na.omit(prosperLoanData$CurrentCreditLines)
CurrentCreditLinesMean <- mean(CurrentCreditLines_na_removed)
CurrentCreditLinesMedian <- median(CurrentCreditLines_na_removed) 

```


Current Credit Lines has a slight positive skew. The mean/median ratio is 
10.31/10. 



```{r echo=FALSE, Univariate_Plots_10, warning=FALSE, message=FALSE}
#using lubridate and extracting the month and year from the loan Origination Date
prosperLoanData$date <- ymd_hms(prosperLoanData$LoanOriginationDate)
prosperLoanData$LoanMonth <- factor(month(prosperLoanData$date))
prosperLoanData$LoanYear <- factor(year(prosperLoanData$date))

ggplot(aes(x=LoanMonth),data=prosperLoanData)+geom_bar()+
  ggtitle("Loan Origination - Monthly")


```


Loans originate mostly during October, December and January

```{r echo=FALSE, Univariate_Plots_11, warning=FALSE, message=FALSE}
ggplot(aes(x=LoanYear),data=prosperLoanData)+geom_bar()+
  ggtitle("Loan Origination - Yearly")
```


During the year 2009, there was a drop. This can be attributed to the economic 
crisis during that year. 

```{r echo=FALSE, Univariate_Plots_12, warning=FALSE, message=FALSE}
table(prosperLoanData$LoanStatus)
```


Most of the loans are either in the completed status or in current status. 


```{r echo=FALSE, Univariate_Plots_23, warning=FALSE, message=FALSE}
summary(prosperLoanData$IncomeRange)
```

Most of the loans are taken by people with income range between 25k USD to 
50k USD and 50k to 75k USD. 

```{r echo=FALSE, Univariate_Plots_13, warning=FALSE, message=FALSE}
summary(prosperLoanData$Term)
table(prosperLoanData$Term)
```


Most of the loans are of duration 36 months (3 years)

```{r echo=FALSE, Univariate_Plots_14, warning=FALSE, message=FALSE}
prosperLoanData_na_omitted <- na.omit(prosperLoanData)
top_10_states <- prosperLoanData_na_omitted  %>% group_by(BorrowerState) %>% 
  summarise(count=n()) %>% arrange(desc(count)) %>% top_n(10)


ggplot(aes(x=BorrowerState, y = count),data=top_10_states)+
  geom_bar(stat="identity")+
  ggtitle("Borrower State Vs count - Bar Chart for the top 10 states")
```


Top 10 states by the number of borrowers were plotted and this the state 
California topped the list. 


```{r echo=FALSE, Univariate_Plots_Occupation, warning=FALSE, message=FALSE}
top_7_occupations <- prosperLoanData_na_omitted  %>% group_by(Occupation) %>% 
  summarise(count=n()) %>% arrange(desc(count)) %>% top_n(7)


ggplot(aes(x=Occupation, y = count),data=top_7_occupations)+
  geom_bar(stat="identity")+
  ggtitle("Occupation Vs count - Bar Chart for the top 7 occupations")
```


Since the most of the loans belonged to the category of "Other", the data is not
fully sufficient to analyse people with what kind of occupation go for the 
highest loans. 


```{r echo=FALSE, Univariate_Plots_HomeBuyer, warning=FALSE, message=FALSE}
print("Borrowers - Are they home owners?")
summary(prosperLoanData$IsBorrowerHomeowner)
```

Almost 50% of the borrowers are home owners


```{r echo=FALSE, Univariate_Plots_Employment_Status, warning=FALSE, message=FALSE}
top_7_occupations <- prosperLoanData_na_omitted  %>% group_by(Occupation) %>% 
  summarise(count=n()) %>% arrange(desc(count)) %>% top_n(7)


ggplot(aes(x=EmploymentStatus),data=prosperLoanData)+
  geom_bar()+
  ggtitle("Employment Status - Histogram")
```


Employment Status data doesnt seem to be complete. Also, as anticipated, the 
employed categroy go for the highest number of loans. 


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.


# Bivariate Analysis

```{r echo=FALSE, Bivariate_Plots_CurrentCreditLines_And_Total_Monthly_Debt}
ggplot(aes(x=CurrentCreditLines, y=MonthlyDebtTotal),data=prosperLoanData)+
  geom_point(alpha=0.5,position="jitter")+xlim(0,30)+ylim(0,10000)
  ggtitle("Current Credit Lines Vs Total Monthly Debt")
  corCoef = cor.test(prosperLoanData$CurrentCreditLines,
                     prosperLoanData$MonthlyDebtTotal,method="pearson")

```


There is a clear positive correlation between Current CreditLines and the Total 
Monthly Debt and the value of the correlation coeffient is 0.47.


```{r echo=FALSE, Bivariate_Plots_CreditScoreRange_And_BorrowerRate}
ggplot(aes(x=CreditScoreRangeLower, y=BorrowerRate),data=prosperLoanData)+
  geom_point(alpha=0.1,position="jitter")+
  ggtitle("Credit Score Range Lower And Borrower Rate")+xlim(450,900)
  corCoef = cor.test(prosperLoanData$CreditScoreRangeLower,
                     prosperLoanData$BorrowerRate,method="pearson")

```


There is a clear negative correlation betwen the CreditScores and the borrower
rate. The correlation coefficient is -0.46. Persons with a lower credit scores
get the loans at a lower borrower rate. 


```{r echo=FALSE, Bivariate_Plots_MonthlyIncome_And_TotalMonthlyDebt}
ggplot(aes(x=StatedMonthlyIncome, y=MonthlyDebtTotal),data=prosperLoanData)+
  geom_point(alpha=0.1,position="jitter")+
  ggtitle("Monthly Income And Total Monthly Debt")+ xlim(0,15000)+ylim(0,6500)
  corCoef = cor.test(prosperLoanData$StatedMonthlyIncome,
                     prosperLoanData$MonthlyDebtTotal,method="pearson")
 
```


There is a positive correlation between the stated Monthly Income and the total
monthly debt. The correlation coefficient is 0.36. 



```{r echo=FALSE, Bivariate_Plots_LoanOriginalAmount_And_BorrowerRate}
ggplot(aes(x=LoanOriginalAmount, y=BorrowerRate),data=prosperLoanData)+
  geom_point(alpha=0.1,position="jitter",size=1)+
  ggtitle("Monthly Income And Total Monthly Debt")#+ xlim(0,15000)+ylim(0,6500)
  corCoef = cor.test(prosperLoanData$LoanOriginalAmount,
                     prosperLoanData$BorrowerRate,method="pearson")
 
```


The correlation between the loan Amount and Borrower Rate is negative. The value 
of the correlation coeffient is -0.32. This would mean that the higher loans are
disbursed at a lower interest rate. 


```{r echo=FALSE, Bivariate_Plots_LoanOriginalAmount_And_CreditScoreRangeLower}
ggplot(aes(x=LoanOriginalAmount, y=CreditScoreRangeLower),data=prosperLoanData)+
  geom_point(alpha=0.1,position="jitter")+
  ggtitle("LoanOriginalAmount And CreditScoreRangeLower")#+ xlim(0,15000)+ylim(0,6500)
  corCoef = cor.test(prosperLoanData$LoanOriginalAmount,
                     prosperLoanData$CreditScoreRangeLower,method="pearson")
 
```

There is a slight positive correlation between loan Amount and Credit Scores. 
Higher the loans, the credit scores can increase. 


```{r echo=FALSE, Bivariate_Plots_StatedMonthlyIncome_And_TotalMontlyDebt}
p1<-ggplot(aes(x=StatedMonthlyIncome, y=MonthlyDebtTotal),data=prosperLoanData)+
  geom_point(alpha=0.025,position="jitter")+
  ggtitle("StatedMonthlyIncome And TotalMontlyDebt")+ xlim(0,15000)+ylim(0,7000)
p1 + facet_wrap(~IsBorrowerHomeowner)
  corCoef = cor.test(prosperLoanData$StatedMonthlyIncome,
                     prosperLoanData$MonthlyDebtTotal,method="pearson")
 
```


Monthly Income Vs Total Monthly Debt was plotted and was facet wrapped by 
IsBorrowerHomeowner.The dispersion is higher in the case of home owners and for 
the non home owners, the dispersion is concentrated more amoung the borrowers 
with an income of around 5000. 


```{r echo=FALSE, Bivariate_Plots_LoanYear_And_LoanOriginalAmount}
ggplot(data=prosperLoanData, aes(x=LoanYear,y=LoanOriginalAmount)) + 
  geom_boxplot()+
  ggtitle("Loan Year And Loan Original Amount")

 
```


The median amount dips in 2009 and has a sharp rise in 2013. 



```{r echo=FALSE, Bivariate_Plots_LoanMonth_And_LoanOriginalAmount}
ggplot(data=prosperLoanData, aes(x=LoanYear,y=LoanOriginalAmount))+ 
  geom_boxplot()+
  ggtitle("Loan Month And Loan Original Amount")


ggplot(data=prosperLoanData, aes(x=LoanMonth, y=LoanOriginalAmount)) + 
  geom_bar(stat='identity')

 
```

Dec to Feb is the period when the loan amounts are usually higher. 

```{r echo=FALSE, Bivariate_Plots_ListingCategory_And_LoanOriginalAmount}
ggplot(data=prosperLoanData, aes(x=ListingCategory..numeric.,
                                 y=LoanOriginalAmount, 
                   fill=ListingCategory..numeric.,)) + 
  geom_boxplot() + scale_fill_discrete(name="Loan Category", labels=cats) +
  ggtitle("Loan Amount and Loan Category")
```

The higher amount loans are taken for Debt consolidation and baby and adoption

```{r echo=FALSE, Bivariate_Plots_IncomeRange_And_LoanOriginalAmount}
ggplot(data=prosperLoanData, aes(x=IncomeRange,y=LoanOriginalAmount, 
                                 color=IncomeRange))+
  geom_boxplot() + scale_x_discrete(labels="") +
  ggtitle("Loan Amount and Income Range")
```

Borrowers with large incomes take large loan.


```{r echo=FALSE, Bivariate_Plots_Employment_Status_And_LoanOriginalAmount}
ggplot(data=subset(prosperLoanData, prosperLoanData$EmploymentStatus!=""), 
       aes(x=EmploymentStatus, y=LoanOriginalAmount, color=EmploymentStatus)) +
  geom_boxplot() + scale_x_discrete(labels="")+ggtitle("Employment Status and 
                                                       Loan Amount")
```

Employed and Self employed are the ones who are taking higher loans


```{r echo=FALSE, Bivariate_Plots_Term_And_LoanOriginalAmount}
ggplot(data=prosperLoanData, aes(x=Term,y=LoanOriginalAmount))+ geom_boxplot()
```

Higher Loans Amounts are repaid thorough longer terms

```{r echo=FALSE, Term_And_BorrowerRate, message=FALSE, warning=FALSE}
ggplot(data=prosperLoanData, aes(x=Term,y=BorrowerRate))+ geom_boxplot()+
  ggtitle("Term and Borrower Rate")
```


Higher the term of repayment does not always means that the interest rates are 
higher. 


```{r echo=FALSE, Term_And_CreditScores, message=FALSE, warning=FALSE}
ggplot(data=prosperLoanData, aes(x=Term,y=CreditScoreRangeLower))+ 
  geom_boxplot() + ggtitle("Term and Credit Scores")+ylim(300,750)
```


Term and Credit Scores - Not much insights can be drawn from this plot. 

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Scores_Amount_BorrowerRate, message=FALSE, warning=FALSE}


ggplot(data=prosperLoanData, 
       aes(x=LoanOriginalAmount, y=BorrowerRate, color=CreditScoreRangeLower)) +
  geom_point(alpha=0.75, position='jitter') + 
  scale_colour_gradient(low="blue", high="red") + 
  ggtitle("Loan Amount by Credit Score and Interest Rate")

```

The top left ones are the ones which takes loans for a smaller amount and at 
higher rates. These are the borrowers with lower credit scores


```{r echo=FALSE, MVP_2, message=FALSE, warning=FALSE}
ggplot(data=subset(prosperLoanData, prosperLoanData$IncomeRange != '$0' & 
                     prosperLoanData$IncomeRange != 'Not displayed'), 
       aes(x=LoanYear, y= MonthlyLoanPayment/StatedMonthlyIncome, 
           fill=IncomeRange)) +geom_boxplot() +ylim(0, 0.1) +
  facet_wrap(~IncomeRange) + theme(axis.text.x = element_text(angle=90))
```



```{r echo=FALSE, MVP_2, message=FALSE, warning=FALSE}
#ggplot(data=subset(prosperLoanData, prosperLoanData$IncomeRange != '$0' & 
#                     prosperLoanData$IncomeRange != 'Not displayed'), 
#       aes(x=LoanYear, y= MonthlyLoanPayment/StatedMonthlyIncome, 
#           fill=IncomeRange)) +geom_boxplot() +ylim(0, 0.1) +
#  facet_wrap(~IncomeRange) + theme(axis.text.x = element_text(angle=90))

ggplot(data=subset(prosperLoanData, prosperLoanData$IncomeRange != '$0' & 
                     prosperLoanData$IncomeRange != 'Not displayed'), 
       aes(x=LoanYear, y= DebtToIncomeRatio, fill=IncomeRange)) +
  geom_boxplot() +ylim(0, .5) +facet_wrap(~IncomeRange)+ 
  theme(axis.text.x = element_text(angle=90))
```


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(data=subset(prosperLoanData, prosperLoanData$CreditScoreRangeLower > 660), aes(x=BorrowerRate, y=LoanOriginalAmount, color=CreditScoreRangeLower)) +geom_point(alpha=0.1, position='jitter') + scale_colour_gradient(low="blue", high="red") +ggtitle("Loan Amount by Credit Score and Interest Rate") +facet_wrap(~LoanYear)
```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data=subset(prosperLoanData, prosperLoanData$DebtToIncomeRatio < .4 ), aes(x=BorrowerRate, y=LoanOriginalAmount, color=DebtToIncomeRatio)) +geom_point(alpha=0.1, position='jitter') +facet_wrap(~LoanYear) + scale_colour_gradient(low="green", high="red") +ggtitle("Loan Amount by DI Ratio and Interest Rate") +xlim( 0, 0.4)

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}
delin <- prosperLoanData[grep("Past Due", prosperLoanData$LoanStatus),] # only delinquent

ggplot(data=subset(delin, delin$IncomeRange != '$0'), aes(y=LoanYear, x=LoanStatus)) +geom_point(aes(color=LoanStatus),size=3, alpha=0.5, position='jitter') +facet_wrap(~IncomeRange) +ggtitle("Delinquent loan exposure summary") +scale_colour_brewer(type="div", palette="Dark2") +scale_x_discrete(labels="")

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!