Exploring Prosper Loan Data by Pyari Singh K
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(lubridate)
library(dplyr)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
prosperLoanData <- read.csv("prosperLoanData.csv")
```



For the explore and summarise data project, the propserLoanData 
data set is what I chose. The data set has 81 columns & 113937 observations.  


```{r}
#converting the ListingCategory to factor variable & associating them 
#for labeling purposes
prosperLoanData$ListingCategory..numeric. <- 
  factor(prosperLoanData$ListingCategory..numeric.)

cats <- c('0 - Not Available', '1 - Debt Consolidation', 
          '2 - Home Improvement', '3 - Business', '4 - Personal Loan', 
          '5 - Student Use', '6 - Auto', '7- Other', '8 - Baby&Adoption', 
          '9 - Boat', '10 - Cosmetic Procedure', '11 - Engagement Ring', 
          '12 - Green Loans', '13 - Household Expenses', 
          '14 - Large Purchases', '15 - Medical/Dental', '16 - Motorcycle', 
          '17 - RV', '18 - Taxes', '19 - Vacation', '20 - Wedding Loans')

#converting Term into a factor variable
prosperLoanData$Term <- factor(prosperLoanData$Term)

```

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots_15, message = FALSE, warning= FALSE}
ggplot(aes(x=LoanOriginalAmount),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Loan Original Amount") +
  labs(x = "Loan Amount ($)", y = "Count")



meanLoanOriginalAmount <- mean(prosperLoanData$LoanOriginalAmount)
medianLoanOriginalAmount <- median(prosperLoanData$LoanOriginalAmount)
```


Loan Original Amount has a positive skew. The mean value for the 
LoanOriginalAmount is 8337 which is greater than the median. The median for the
LoanOriginalAmount is 6500. 

```{r echo=FALSE, Univariate_Plots_16, message = FALSE, warning= FALSE}
ggplot(aes(x=MonthlyLoanPayment),data=prosperLoanData)+
  geom_histogram(binwidth = 50)+
  ggtitle("Histogram for Monthly Loan Payment")+
  coord_cartesian(xlim=c(0:1500)) +
  labs(x = "Monthly Loan Payment ($)", y = "Count")


monthlyLoanPaymentMean <- mean(prosperLoanData$MonthlyLoanPayment)
monthlyLoanPaymentMedian <- median(prosperLoanData$MonthlyLoanPayment) 
```

Monthly Loan Payment has a positive skew. The median is less than the mean for 
this data. The median for the monthly Loan Payment is 217 and the mean is 272. 


```{r echo=FALSE, Univariate_Plots_17, message = FALSE, warning= FALSE}
ggplot(aes(x=CreditScoreRangeLower),data=prosperLoanData)+
  geom_histogram(binwidth = 10)+
  ggtitle("Histogram for Credit Score Range Lower") + xlim(400, 900) +
  labs(x = "Credit Score Range (Lower)", y = "Count")



creditScoreRangeLower_na_removed<-na.omit(prosperLoanData$CreditScoreRangeLower)
creditScoreRangeLowerMean <- mean(creditScoreRangeLower_na_removed)
creditScoreRangeLowerMedian <- median(creditScoreRangeLower_na_removed) 
```


credit score range lower has a normal distribution. The mean/median in this case 
is 685.56/680


```{r echo=FALSE, Univariate_Plots_18, message = FALSE, warning= FALSE}
ggplot(aes(x=BorrowerRate),data=prosperLoanData)+
  geom_histogram(binwidth = 0.5)+
  ggtitle("Histogram for Borrower Rate") +
  labs(x = "Borrower Rate", y = "Count")


BorrowerRateMean <- mean(prosperLoanData$BorrowerRate)
BorrowerRateMedian <- median(prosperLoanData$BorrowerRate) 
```


Borrower Rate also has a normal distribution and the mean/med ratio is 0.19/0.18
However, there are various spikes especially between 0.24 and 0.31.


```{r echo=FALSE, Univariate_Plots_19, message=FALSE, warning=FALSE}
ggplot(aes(x=DebtToIncomeRatio),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Debt to Income Ratio")+xlim(0,1.75) +
  labs(x = "Debt to income ratio", y = "Count")


DIRatio_na_removed<- na.omit(prosperLoanData$DebtToIncomeRatio)
DIRatioMean <- mean(DIRatio_na_removed)
DIRatioMedian <- median(DIRatio_na_removed) 
```


Debt To Income Ratio exhibits a slight positive skew with mean > median. 
Mean is 0.275 and median is 0.22. DI ratio 
is less than 0.5 for a good percentage of pepole. 



```{r echo=FALSE, Univariate_Plots_20, warning=FALSE, message=FALSE}
ggplot(aes(x=StatedMonthlyIncome),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Stated Monthly Income")+xlim(0,20000) +
  labs(x = "Staged monthly income ($)", y = "Count")


StatedMonthlyIncomeMean <- mean(prosperLoanData$StatedMonthlyIncome)
StatedMonthlyIncomeMedian <- median(prosperLoanData$StatedMonthlyIncome) 
```


Stated Monthly Income has a positive skew with mean > median. Mean for the 
Stated Monthly Income is 5608 and the median is 4666. 


```{r echo=FALSE, Univariate_Plots_21, warning=FALSE, message=FALSE}

prosperLoanData$MonthlyDebtTotal <- prosperLoanData$DebtToIncomeRatio * 
  prosperLoanData$StatedMonthlyIncome

ggplot(aes(x=MonthlyDebtTotal),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Total Monthly Debt")+xlim(0,7500) +
  labs(x = "Monthly Debt Total ", y = "Count")


TotalMonthlyDebt_na_removed<- na.omit(prosperLoanData$MonthlyDebtTotal)
TotalMonthlyDebtMean <- mean(TotalMonthlyDebt_na_removed)
TotalMonthlyDebtMedian <- median(TotalMonthlyDebt_na_removed) 

summary(prosperLoanData$MonthlyDebtTotal)
```


MonthlyDebtTotal has been added as a new column. This was derived by multiplying
the debt to Income Ratio to the stated Monthly Income. 

Total Monthly debt has a positive skew on its histogram and the mean is greater 
than the median. The mean is 1224. The median is 1061. The first quartile for 
this data is 637 and the third quartile is 1622. 

```{r echo=FALSE, Univariate_Plots_22, warning=FALSE, message=FALSE}


ggplot(aes(x=EmploymentStatusDuration),data=prosperLoanData)+geom_histogram()+
  ggtitle("Histogram for Employment Status Duration")+xlim(0,700) +
  labs(x = "Employment status duration (Days)", y = "Count")


EmploymentStatusDuration_na_removed<- 
  na.omit(prosperLoanData$EmploymentStatusDuration)

EmploymentStatusDurationMean <- mean(EmploymentStatusDuration_na_removed)
EmploymentStatusDurationMedian <- median(EmploymentStatusDuration_na_removed) 

```


The Employment Status Duration exhibits a positive skew. The mean of the data is
higher than the median. The mean is 96 and the median is 67. It is observed that
more number of employees worked for shorter durations. 



```{r echo=FALSE, Univariate_Plots_9, warning=FALSE, message=FALSE}
ggplot(aes(x=CurrentCreditLines),data=prosperLoanData)+
  geom_histogram(binwidth = 1)+
  ggtitle("Histogram for Current Credit Lines")+xlim(0,70) +
  labs(x = "Current credit lines", y = "Count")


CurrentCreditLines_na_removed<- na.omit(prosperLoanData$CurrentCreditLines)

CurrentCreditLinesMean <- mean(CurrentCreditLines_na_removed)
CurrentCreditLinesMedian <- median(CurrentCreditLines_na_removed) 

```


Current Credit Lines has a slight positive skew. The mean/median ratio is 
10.31/10. 



```{r echo=FALSE, Univariate_Plots_10, warning=FALSE, message=FALSE}
#using lubridate to extract the month & year from the loan Origination Date
prosperLoanData$date <- ymd_hms(prosperLoanData$LoanOriginationDate)

prosperLoanData$LoanMonth <- factor(month(prosperLoanData$date))
prosperLoanData$LoanYear <- factor(year(prosperLoanData$date))

ggplot(aes(x=LoanMonth),data=prosperLoanData)+geom_bar()+
  ggtitle("Loan Origination - Monthly") +
  labs(x = "Month", y = "Count")



```


Loans originate mostly during October, December and January

```{r echo=FALSE, Univariate_Plots_11, warning=FALSE, message=FALSE}
ggplot(aes(x=LoanYear),data=prosperLoanData)+geom_bar()+
  ggtitle("Loan Origination - Yearly") +
  labs(x = "Year", y = "Count")

```


During the year 2009, there was a drop. This can be attributed to the economic 
crisis during that year. 

```{r echo=FALSE, Univariate_Plots_12, warning=FALSE, message=FALSE}
table(prosperLoanData$LoanStatus)
```


Most of the loans are either in the completed status or in current status. 


```{r echo=FALSE, Univariate_Plots_23, warning=FALSE, message=FALSE}
summary(prosperLoanData$IncomeRange)
```

Most of the loans are taken by people with income range between 25k USD to 
50k USD and 50k to 75k USD. 

```{r echo=FALSE, Univariate_Plots_13, warning=FALSE, message=FALSE}
summary(prosperLoanData$Term)
table(prosperLoanData$Term)
```


Most of the loans are of duration 36 months (3 years)

```{r echo=FALSE, Univariate_Plots_14, warning=FALSE, message=FALSE}
prosperLoanData_na_omitted <- na.omit(prosperLoanData)
top_10_states <- prosperLoanData_na_omitted  %>% group_by(BorrowerState) %>% 
  summarise(count=n()) %>% arrange(desc(count)) %>% top_n(10)


ggplot(aes(x=BorrowerState, y = count),data=top_10_states)+
  geom_bar(stat="identity")+
  ggtitle("Borrower State Vs count - Bar Chart for the top 10 states") +
  labs(x = "State (US)", y = "Count")

```


Top 10 states by the number of borrowers were plotted and this the state 
California topped the list. 


```{r echo=FALSE, Univariate_Plots_Occupation, warning=FALSE, message=FALSE}
top_7_occupations <- prosperLoanData_na_omitted  %>% group_by(Occupation) %>% 
  summarise(count=n()) %>% arrange(desc(count)) %>% top_n(7)


ggplot(aes(x=Occupation, y = count),data=top_7_occupations)+
  geom_bar(stat="identity")+
  ggtitle("Occupation Vs count - Bar Chart for the top 7 occupations") +
  labs(x = "Borrower Occupation", y = "Count")

```


Since the most of the loans belonged to the category of "Other", the data is not
fully sufficient to analyse people with what kind of occupation go for the 
highest loans. 


```{r echo=FALSE, Univariate_Plots_HomeBuyer, warning=FALSE, message=FALSE}
print("Borrowers - Are they home owners?")
summary(prosperLoanData$IsBorrowerHomeowner)
```

Almost 50% of the borrowers are home owners


```{r echo=FALSE, UVP_Employment_Status, warning=FALSE, message=FALSE}
top_7_occupations <- prosperLoanData_na_omitted  %>% group_by(Occupation) %>% 
  summarise(count=n()) %>% arrange(desc(count)) %>% top_n(7)


ggplot(aes(x=EmploymentStatus),data=prosperLoanData)+
  geom_bar()+
  ggtitle("Employment Status - Histogram") +
  labs(x = "Employment Status", y = "Count")

```


Employment Status data doesnt seem to be complete. Also, as anticipated, the 
employed categroy go for the highest number of loans. 


# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
The data set has around 113937 observations and around 81 columns. For our 
analysis a combination of discrete and continuous variables are being used. All 
of the variables are not being used. We are using around 20 variables here. 

### What is/are the main feature(s) of interest in your dataset?
The data set is huge and the more we explore, the more can we get deeper into 
the data. For now, the focus is on some of the below areas, which was found 
interesting and intriguing:
credit score, debt to income ratio, income, loan amount, interest rate, term etc

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
Loan Category, Loan Amount, Loan Status etc are available and these will help 
analyse the data better

### Did you create any new variables from existing variables in the dataset?
LoanYear, LoanMonth etc were derived from the LoanOriginationDate. Also, a 
new column MonthlyDebtTotal was added by multiplying DI ratio with the stated 
Monthly Income

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
Term and ListingCategory..numeric. were converted to factor variables. They were 
initially categorical values. It was done while making a bar plot 
which expected that these values are discrete. 

# Bivariate Plots Section


# Bivariate Analysis

```{r echo=FALSE, Bivariate_Plots_CurrentCreditLines_And_Total_Monthly_Debt, warning=FALSE, message=FALSE}
ggplot(aes(x=CurrentCreditLines, y=MonthlyDebtTotal),data=prosperLoanData) +
  labs(x = "Current Credit Lines", y = "Monthly Debt Total") +
  geom_point(alpha=0.5,position="jitter")+xlim(0,30)+ylim(0,10000)
  ggtitle("Current Credit Lines Vs Total Monthly Debt") 
  corCoef = cor.test(prosperLoanData$CurrentCreditLines,
                     prosperLoanData$MonthlyDebtTotal,method="pearson")

```


There is a clear positive correlation between Current CreditLines and the Total 
Monthly Debt and the value of the correlation coeffient is 0.47.


```{r echo=FALSE, Bivariate_Plots_CreditScoreRange_And_BorrowerRate, warning=FALSE, message=FALSE}
ggplot(aes(x=CreditScoreRangeLower, y=BorrowerRate),data=prosperLoanData)+
  labs(x = "Credit Score Range (Lower)", y = "Borrower Rate") +
  geom_point(alpha=0.1,position="jitter")+
  ggtitle("Credit Score Range Lower And Borrower Rate")+xlim(450,900)
  corCoef = cor.test(prosperLoanData$CreditScoreRangeLower,
                     prosperLoanData$BorrowerRate,method="pearson")

```


There is a clear negative correlation betwen the CreditScores and the borrower
rate. The correlation coefficient is -0.46. Persons with a lower credit scores
get the loans at a lower borrower rate. 


```{r echo=FALSE, Bivariate_Plots_MonthlyIncome_And_TotalMonthlyDebt, warning=FALSE, message=FALSE}
ggplot(aes(x=StatedMonthlyIncome, y=MonthlyDebtTotal),data=prosperLoanData)+
  labs(x = "Stated Monthly Income", y = "Monthly Debt Total") +
  geom_point(alpha=0.1,position="jitter")+
  ggtitle("Monthly Income And Total Monthly Debt")+ xlim(0,15000)+ylim(0,6500)
  corCoef = cor.test(prosperLoanData$StatedMonthlyIncome,
                     prosperLoanData$MonthlyDebtTotal,method="pearson")
 
```


There is a positive correlation between the stated Monthly Income and the total
monthly debt. The correlation coefficient is 0.36. 



```{r echo=FALSE, Bivariate_Plots_LoanOriginalAmount_And_BorrowerRate, warning=FALSE, message=FALSE}
ggplot(aes(x=LoanOriginalAmount, y=BorrowerRate),data=prosperLoanData)+
  labs(x = "Original Loan Amount", y = "Borrower Rate") +
  geom_point(alpha=0.1,position="jitter",size=1)+
  ggtitle("Monthly Income And Total Monthly Debt")#+ xlim(0,15000)+ylim(0,6500)
  corCoef = cor.test(prosperLoanData$LoanOriginalAmount,
                     prosperLoanData$BorrowerRate,method="pearson")
 
```


The correlation between the loan Amount and Borrower Rate is negative. The value 
of the correlation coeffient is -0.32. This would mean that the higher loans are
disbursed at a lower interest rate. 


```{r echo=FALSE, Bivariate_Plots_LoanOriginalAmount_And_CreditScoreRangeLower, warning=FALSE, message=FALSE}
ggplot(aes(x=LoanOriginalAmount, y=CreditScoreRangeLower),data=prosperLoanData)+
  labs(x = "Original Loan Amount", y = "Credit Score Range (Lower)") +
  geom_point(alpha=0.1,position="jitter")+
  ggtitle("LoanOriginalAmount And CreditScoreRangeLower")#+ xlim(0,15000)+ylim(0,6500)
  corCoef = cor.test(prosperLoanData$LoanOriginalAmount,
                     prosperLoanData$CreditScoreRangeLower,method="pearson")
 
```

There is a slight positive correlation between loan Amount and Credit Scores. 
Higher the loans, the credit scores can increase. 


```{r echo=FALSE, Bivariate_Plots_StatedMonthlyIncome_And_TotalMontlyDebt, warning=FALSE, message=FALSE}
p1<-ggplot(aes(x=StatedMonthlyIncome, y=MonthlyDebtTotal),data=prosperLoanData)+
  labs(x = "Stated Monthly Income", y = "Total Monthly Debt") +
  geom_point(alpha=0.025,position="jitter")+
  ggtitle("StatedMonthlyIncome And TotalMontlyDebt")+ xlim(0,15000)+ylim(0,7000)
p1 + facet_wrap(~IsBorrowerHomeowner)
  corCoef = cor.test(prosperLoanData$StatedMonthlyIncome,
                     prosperLoanData$MonthlyDebtTotal,method="pearson")
 
```


Monthly Income Vs Total Monthly Debt was plotted and was facet wrapped by 
IsBorrowerHomeowner.The dispersion is higher in the case of home owners and for 
the non home owners, the dispersion is concentrated more amoung the borrowers 
with an income of around 5000. 


```{r echo=FALSE, Bivariate_Plots_LoanYear_And_LoanOriginalAmount, warning=FALSE, message=FALSE}
ggplot(data=prosperLoanData, aes(x=LoanYear,y=LoanOriginalAmount)) + 
  labs(x = "Year", y = "Original Loan Amount") +
  geom_boxplot()+
  ggtitle("Loan Year And Loan Original Amount")

 
```


The median amount dips in 2009 and has a sharp rise in 2013. 



```{r echo=FALSE, Bivariate_Plots_LoanMonth_And_LoanOriginalAmount, warning=FALSE, message=FALSE}
ggplot(data=prosperLoanData, aes(x=LoanYear,y=LoanOriginalAmount))+ 
  labs(x = "Year", y = "Original Loan Amount") +
  geom_boxplot()+
  ggtitle("Loan Month And Loan Original Amount")


ggplot(data=prosperLoanData, aes(x=LoanMonth, y=LoanOriginalAmount)) + 
  labs(x = "Month", y = "Original Loan Amount") +
  geom_bar(stat='identity')

 
```

Dec to Feb is the period when the loan amounts are usually higher. 

```{r echo=FALSE, Bivariate_Plots_ListingCategory_And_LoanOriginalAmount, fig.width=15, fig.height=12.5, warning=FALSE, message=FALSE}
ggplot(data=prosperLoanData, aes(x=ListingCategory..numeric.,
                                 y=LoanOriginalAmount, 
                   fill=ListingCategory..numeric.)) + 
  labs(x = "Listing Category", y = "Original Loan Amount") +
  geom_boxplot() + scale_fill_discrete(name="Loan Category", labels=cats) +
  ggtitle("Loan Amount and Loan Category")
```

The higher amount loans are taken for Debt consolidation and baby and adoption

```{r echo=FALSE, Bivariate_Plots_IncomeRange_And_LoanOriginalAmount, warning=FALSE, message=FALSE}
ggplot(data=prosperLoanData, aes(x=IncomeRange,y=LoanOriginalAmount, 
                                 color=IncomeRange))+
  labs(x = "Income Range", y = "Original Loan Amount") +
  geom_boxplot() + scale_x_discrete(labels="") +
  ggtitle("Loan Amount and Income Range")
```

Borrowers with large incomes take large loan.


```{r echo=FALSE, Bivariate_Plots_Employment_Status_And_LoanOriginalAmount, warning=FALSE, message=FALSE}
ggplot(data=subset(prosperLoanData, prosperLoanData$EmploymentStatus!=""), 
       aes(x=EmploymentStatus, y=LoanOriginalAmount, color=EmploymentStatus)) +
  geom_boxplot() + scale_x_discrete(labels="")+ggtitle("Employment Status and Loan Amount") +
  labs(x = "Employment Status", y = "Original Loan Amount") 


```

Employed and Self employed are the ones who are taking higher loans


```{r echo=FALSE, Bivariate_Plots_Term_And_LoanOriginalAmount, warning=FALSE, message=FALSE}
ggplot(data=prosperLoanData, aes(x=Term,y=LoanOriginalAmount))+ geom_boxplot() +
labs(x = "Term (Months)", y = "Original Loan Amount") 
```

Higher Loans Amounts are repaid thorough longer terms

```{r echo=FALSE, Term_And_BorrowerRate, message=FALSE, warning=FALSE}
ggplot(data=prosperLoanData, aes(x=Term,y=BorrowerRate))+ geom_boxplot()+
  labs(x = "Term (Months)", y = "Borrower Rate") +
  ggtitle("Term and Borrower Rate")
```


Higher the term of repayment does not always means that the interest rates are 
higher. 


```{r echo=FALSE, Term_And_CreditScores, message=FALSE, warning=FALSE}
ggplot(data=prosperLoanData, aes(x=Term,y=CreditScoreRangeLower))+ 
  geom_boxplot() + ggtitle("Term and Credit Scores")+ylim(300,750) +
    labs(x = "Term (Months)", y = "Credit Score Range (Lower)") 

```


Term and Credit Scores - Not much insights can be drawn from this plot. 

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
Total Monthly Debt and Current Credit Lines have a strong positive correlation
Longer the term of the loan, it is found that the loan amount is also larger 
Employed and self employed ones borrow larger amounts 

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
Could not do this analysis. The credit scores and the salary ranges, credit 
scores and employment status are some of the relationships I wanted to analyse.

### What was the strongest relationship you found?
CurrentCreditLines and MonthlyDebtTotal had a strong +ve correlation with a 
correlation coefficient of 0.47. Also, credit scores and borrower rate have a 
strong positive correlation with a correlation coefficient of 0.46


# Multivariate Plots Section

```{r echo=FALSE, Scores_Amount_BorrowerRate, message=FALSE, warning=FALSE, fig.width=16, fig.height=12.5}


ggplot(data=prosperLoanData, 
       aes(x=LoanOriginalAmount, y=BorrowerRate, color=CreditScoreRangeLower)) +
  geom_jitter(alpha=0.15, position='jitter') + 
  scale_colour_gradient(low="blue", high="red") + 
  ggtitle("Loan Amount by Credit Score and Interest Rate") +
  labs(x = "Loan Original Amount", y = "Borrower Rate") 


```

The top left ones are the ones which takes loans for a smaller amount and at 
higher rates. These are the borrowers with lower credit scores


```{r echo=FALSE, MVP_2, message=FALSE, warning=FALSE}
ggplot(data=subset(prosperLoanData, prosperLoanData$IncomeRange != '$0' & 
      prosperLoanData$IncomeRange != 'Not displayed'), 
      aes(x=LoanYear, y= MonthlyLoanPayment/StatedMonthlyIncome, 
      fill=IncomeRange)) +geom_boxplot() +ylim(0, 0.1) +
      facet_wrap(~IncomeRange) + theme(axis.text.x = element_text(angle=90))+
      ggtitle("Loan Year - Debt coverage - Facted by Income Range") +
      labs(x = "Year", y = "Monthly Loan Payments / Stated Monthly Income") 
```

MonthlyLoanPayment was divided by the StatedMonthlyIncome to observe the 
yearwise debt coverage ratio. This was faceted by IncomeRange. 
All the income ranges have an uptick in the most recent years.

```{r echo=FALSE, MVP_3, message=FALSE, warning=FALSE}
ggplot(data=subset(prosperLoanData, prosperLoanData$IncomeRange != '$0' & 
      prosperLoanData$IncomeRange != 'Not displayed'), 
      aes(x=LoanYear, y= DebtToIncomeRatio, fill=IncomeRange)) +
      geom_boxplot() +ylim(0, .5) +facet_wrap(~IncomeRange)+ 
      theme(axis.text.x = element_text(angle=90)) +
      labs(x = "Year", y = "Debt To Income Ratio") 

```


Most of the borrowers seem to have DI ratio of 0.2 to 0.3 and is showing an 
uptick in the most recent years. Borrowers with income range $100k + have a
low DI ratio and is around 0.15-0.20 

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Borrowers with lower credit scores takes loans for a smaller amount, however at 
a higher interest rate

Also, it was observed that borrowers with the highest income have a lower 
percentage of debt. 

### Were there any interesting or surprising interactions between features?
Taking a look at the DI ratio, it was observed that higher the income, 
the lower is the percentage of debt. 

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One, message=FALSE,warning=FALSE}
ggplot(data=subset(prosperLoanData, prosperLoanData$CreditScoreRangeLower > 660), aes(x=BorrowerRate, y=LoanOriginalAmount, color=CreditScoreRangeLower)) +geom_point(alpha=0.1, position='jitter') + scale_colour_gradient(low="blue", high="red") +ggtitle("Loan Amount by Credit Score and Interest Rate") +facet_wrap(~LoanYear)
```

### Description One
The above plot shows how the lending platform Prosper has matured over years. 
Red shows higher credit scores and we can see that by the year 2014, blue 
dominates,which would also mean that the over all risk exposure has increased. 
(For better visualisation, only credit scores from the first quartile are 
plotted here)

### Plot Two
```{r echo=FALSE, Plot_Two, message=FALSE, warning=FALSE}
ggplot(data=subset(prosperLoanData, prosperLoanData$DebtToIncomeRatio < .4 ), 
       aes(x=BorrowerRate, y=LoanOriginalAmount, 
           color=DebtToIncomeRatio)) +geom_point(alpha=0.1, 
                                  position='jitter') +facet_wrap(~LoanYear) + 
      scale_colour_gradient(low="green", high="red") +
      ggtitle("Loan Amount by DI Ratio and Interest Rate") +xlim( 0, 0.4)



```

### Description Two

As the prosper lending platform matures, probably the risk management also has
improved. This could be the reason, the borrower rates are decreasing for larger 
loans, mainly for those with lower debt to income ratio.  

### Plot Three
```{r echo=FALSE, Plot_Three}
delin <- prosperLoanData[grep("Past Due", prosperLoanData$LoanStatus),] 
# only delinquent

ggplot(data=subset(delin, delin$IncomeRange != '$0'), 
       aes(y=LoanYear, x=LoanStatus)) +geom_point(aes(color=LoanStatus),size=3, 
      alpha=0.5, position='jitter') +
  facet_wrap(~IncomeRange) +ggtitle("Delinquent loan exposure 
                        summary") +scale_colour_brewer(type="div", 
                                  palette="Dark2") +scale_x_discrete(labels="")

```

### Description Three
Only deliquent loans were subsetted here. The majority of the delinquencies are 
clustered in 1-15days past due. Maximum delinqueny is happening for the 
borrowers having income ranges between $25k-50k and $50-75k.

------

# Reflection
The data set has almost 114,000 loans. 

The most time consuming part of this analysis was to find some variables which 
will help make some good analysis & to list down what are the analysis that 
has to be performed. 

There was too much of information that, it was not difficult to atleast perform 
the single variable and bi variable analysis. There were too many choices. 
However, drilling down into multi variable analysis from the results of the bi 
variable analysis was again difficult. 

Making the choice for plotting three plots as the outcome of the entire analysis 
was again difficult since there was too much of data to be chosen from.    

Regression models & some predictions are what I would have tried working on if 
time permitted.  For example, preclosure, delinquency, late payment etc if it 
can be predicted, it would have been good. 

One of the good decisions was to restrict the analysis to limited variables and 
not look into the entire data to perform the analysis. This helped finish the 
analysis comparitively faster. 

# References
https://rpubs.com/yugoli/UdacityP3


